---
title: "Practical Machine Learning"
author: "Regina Hong"
date: "27 May 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

## What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

Your submission for the Peer Review portion should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-). Apply your machine learning algorithm to the 20 test cases available in the test data above and submit your predictions in appropriate format to the Course Project Prediction Quiz for automated grading.

## Choosing the Prediction Model

Two models will be tested using decision tree and random forest algorithms. The model with the highest accuracy will be chosen as our final model.

```{r}
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
```

## Load Data

#### To remove the "#DIV/0!" data and replace with NA value.

```{r}
trainingset <- read.csv("D:/R working file/Course 8 Practical Machine Learning/Week 4/pml-training.csv",na.strings = c("NA","#DIV/0!","")) 
testingset <- read.csv("D:/R working file/Course 8 Practical Machine Learning/Week 4/pml-testing.csv",na.strings = c("NA","#DIV/0!",""))
```

#### To check dimensions for number of variables and number of observations

```{r}
dim(trainingset)
dim(testingset)
```

#### To delete columns with all missing values

```{r}
trainingset<-trainingset[,colSums(is.na(trainingset)) == 0]
testingset <-testingset[,colSums(is.na(testingset)) == 0]
```

#### To delete those irrelevant variables

```{r}
trainingset   <-trainingset[,-c(1:7)]
testingset <-testingset[,-c(1:7)]
```

#### To look at the new datasets

```{r}
dim(trainingset)
dim(testingset)
head(trainingset)
head(testingset)
```

## Partitioning Training Data Set

In order to perform cross-validation, the training data set is partionned into 2 sets: subTraining (75%) and subTest (25%).
This will be performed using random subsampling without replacement.

```{r}
subsamples <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
subTraining <- trainingset[subsamples, ] 
subTesting <- trainingset[-subsamples, ]
dim(subTraining)
dim(subTesting)
head(subTraining)
head(subTesting)
```

## Plot a Bar Plot

```{r}
plot(subTraining$classe, col="blue", main="Bar Plot of levels of the variable classe within the subTraining data set", xlab="classe levels", ylab="Frequency")
```

Level A is the most frequent occurrences while level D is the least frequent occurrences.

## First Prediction : Using Decision Tree

```{r}
model1 <- rpart(classe ~ ., data=subTraining, method="class")
prediction1 <- predict(model1, subTesting, type = "class")
rpart.plot(model1, main="Classification Tree", extra=102, under=TRUE, faclen=0)
confusionMatrix(prediction1, subTesting$classe)
```

## Second Prediction : Using Random Forest

```{r}
model2 <- randomForest(classe ~. , data=subTraining, method="class")
prediction2 <- predict(model2, subTesting, type = "class")
confusionMatrix(prediction2, subTesting$classe)
```

## Decision

Random Forest algorithm performed better than Decision Trees.Accuracy for Random Forest model was 0.991 (95% CI: (0.9884, 0.9938)) compared to 0.664 (95% CI: (0.6509, 0.6776)) for Decision Tree model. The random Forest model is choosen. The accuracy of the model is 0.991. The expected out-of-sample error is estimated at 0.005, or 0.5%. The expected out-of-sample error is calculated as 1 - accuracy for predictions made against the cross-validation set. Our Test data set comprises 20 cases. With an accuracy above 99% on our cross-validation data, we can expect that very few, or none, of the test samples will be missclassified.
